---
title: "README"
output:
  md_document:
    variant: gfm
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path = "man/figures/README-"
)
```

# loadflex: Models and Tools for Watershed Flux Estimates

The `loadflex` R package implements several of the most common methods for modeling and
predicting watershed solute fluxes and concentrations, including interpolation
and regression models, period-weighted averaging, and the composite method.
`loadflex` integrates seamlessly with the USGS's
[rloadest](https://code.usgs.gov/water/analysis-tools/rloadest) package and with native R
regression models. It offers a uniform interface for any model type, with which
you can quickly fit models, generate predictions, and aggregate to monthly or
annual values.

To see the recommended citation for this package, please run `citation('loadflex')` at the R prompt.
```{r eval=TRUE}
citation('loadflex')
```

## Installation

To install the `loadflex` package, use the `remotes` package as follows (running `install.packages('remotes')` first if needed). To use `remotes::install_gitlab()` it is convenient to set a [GitLab Personal Access Token (PAT)](https://docs.gitlab.com/ee/user/profile/personal_access_tokens.html). Similarly, to use `remotes::install_github()` it is convenient to set a [GitHub PAT](https://docs.github.com/en/authentication/keeping-your-account-and-data-secure/managing-your-personal-access-tokens). There are [several methods](https://usethis.r-lib.org/articles/git-credentials.html) for setting your PATs within R; the simplest is to call `Sys.setenv(GITLAB_PAT="xxxx", GITHUB_PAT="yyyy"),
replacing xxxx and yyyy with the PATs you established on the GitLab and GitHub websites.

You will also need a compiler to install `smwrStats`, `smwrQW`, and `rloadest` -- for Windows, see https://cran.r-project.org/bin/windows/Rtools/rtools43/rtools.html. For Mac, see https://mac.r-project.org/tools/.

```{r eval=FALSE}
# install.packages(remotes) # run if needed
remotes::install_gitlab("water/analysis-tools/smwrData", host = "code.usgs.gov")
remotes::install_gitlab("water/analysis-tools/smwrBase", host = "code.usgs.gov")
remotes::install_gitlab("water/analysis-tools/smwrGraphs", host = "code.usgs.gov")
remotes::install_gitlab("water/analysis-tools/smwrStats", host = "code.usgs.gov") # needs compilation
remotes::install_gitlab("water/analysis-tools/smwrQW", host = "code.usgs.gov")    # needs compilation
remotes::install_github("USGS-R/rloadest")  # needs compilation
remotes::install_github("appling/unitted")
remotes::install_github("DOI-USGS/EGRET")
remotes::install_github("DOI-USGS/loadflex")
```

Also please see the installation FAQ on the wiki
(https://github.com/DOI-USGS/loadflex/wiki/Installation-FAQ) if you run into trouble.


## Getting Started

To get started, load the package with `library(loadflex)` and type `?loadflex` or `vignette('intro_to_loadflex')`.


## Development and Maintenance Status

Project funding has ended and our maintenance time is limited,
but we do attempt to provide bug fixes and lightweight support
as we are able. Submit questions or suggestions to
[https://github.com/DOI-USGS/loadflex/issues](https://github.com/DOI-USGS/loadflex/issues).


## Contributing

We want to encourage a warm, welcoming, and safe environment for contributing to this project.
See [CODE_OF_CONDUCT.md](https://github.com/DOI-USGS/loadflex/blob/main/CODE_OF_CONDUCT.md) for more information.

For technical details on how to contribute,
see [CONTRIBUTING.md](https://github.com/DOI-USGS/loadflex/blob/main/CONTRIBUTING.md)


## Contributing

We want to encourage a warm, welcoming, and safe environment for contributing to this project. See [CODE_OF_CONDUCT.md](https://github.com/DOI-USGS/loadflex/blob/main/CODE_OF_CONDUCT.md) for more information.

For technical details on how to contribute, see [CONTRIBUTING.md](https://github.com/DOI-USGS/loadflex/blob/main/CONTRIBUTING.md)


## Development History

`loadflex` was created 2013-2015 by [Alison Appling](https://github.com/aappling-usgs) 
and [Miguel Leon](https://github.com/miguelcleon) with the support of [Bill 
McDowell](https://colsa.unh.edu/faculty/mcdowell) and the [McDowell 
lab](http://wrrc.unh.edu/mcdowell-lab-current) at the University of New 
Hampshire. Funding for the project was provided by the National 
Science Foundation, USDA National Institute of Food and Agriculture, and the NH 
Agricultural Experiment Station. Funding for the example dataset from the 
Lamprey River was provied by the EPA, NH Water Resources Research Center, NH 
Agricultural Experiment Station, NH Sea Grant, USGS, and NSF.

Additional development and maintenance in 2016-2017 was done by 
[Alison Appling](https://github.com/aappling-usgs),
[Lindsay Platt](https://github.com/lindsayplatt), and
[David Watkins](https://github.com/wdwatkins) with support from the USGS National Water Quality Program and the USGS Office of Water Information.


## Model Archive

The following version of R and package dependencies were used most recently to pass the
embedded tests within this package. There is no guarantee of reproducible results using
future versions of R or updated versions of package dependencies; however, we aim to test
and update future modeling environments.

```{r echo=FALSE}
suppressPackageStartupMessages(library(loadflex))
info_txt <- capture.output(sessioninfo::session_info())
clean_txt <- info_txt[!grepl("(^ \\[)|(^ pandoc)|(^ quarto)", info_txt)]
cat(clean_txt, sep = '\n')
```


```{r disclaimer, child="DISCLAIMER.md", eval=TRUE}
```
